---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Lena Parnassa"
date: today

format:
  html:
    toc: true
    embed-resources: true

execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---
::: {.callout-tip icon=false}

## Github Repo Link
[https://github.com/stat301-2-2024-winter/final-project-2-lena-parnassa](https://github.com/stat301-2-2024-winter/final-project-2-lena-parnassa)

:::
```{r , echo=FALSE}
#load libraries
library(tidyverse)
library(tidymodels)
library(here)
library(knitr)

# handle common conflicts
tidymodels_prefer()

#load data
load(here("results/fl_split.rda"))
load(here("results/basic_recipe.rda"))
load(here("results/fit_lm_1.rda"))
load(here("results/fit_enet_1.rda"))
load(here("results/fit_knn_1.rda"))
```


## Data/Project Overview:
I'm looking at data from [The Echinacea Project](https://echinaceaproject.org/) at the Chicago Botanic Garden, which is the lab I'm working in. Specifically, I'm working with our [core dataset](https://echinaceaproject.org/datasets/expt1-core/), and a datset on [insect damage](https://echinaceaproject.org/datasets/km-aphid-abundance-and-plant-damage-2004-2012/). My goal is to predict the achene count. Achenes are small fruits produced by echinacea plants, and a healthy flowering head can easily produce upwards of 100 achenes, which makes counting them a time and labor intensive process. I filtered my datasets to only include plants with flowering heads in a given year, as the alternative (basal or dead plants) are incapable of producing achenes. My cleaned dataset has 2,865 observations of 42 variables.


## Analysis Plan

### Assesment Metric
I plan on using RMSE as my assessment metric because my outcome variable has a lot of large outliers and RMSE will be better at handling them than $R^2$ or MAE.  

### Data Splitting
I split my data with a proportion of 0.7 because its on the smaller side. I did a v-fold cross validation with 5 folds and 3 repeats. While I know that is fewer than would be most appropriate for a dataset of this size, my computer is old and probably not capable of handling much more.

### Model Types
I'm using a linear regression model as my baseline model, and also using elastic net, KNN, random forest, and boosted tree models. Each model will be tested on two recipes, one "kitchen sink" recipe that I have already created, and a second recipe with more feature engineering that will be based on the performance of the first recipe. I have designed parametric and non-parametric versions of my first recipe and will do the same when creating the second.

## Fit Demonstration
So far I've defined and fitted my linear regression and KNN models using my kitchen sink recipes:
```{r , echo=FALSE}
model_results <- as_workflow_set(
  lm_basic = fit_lm_1 ,
  knn_basic = fit_knn_1 
)

model_results |>
  collect_metrics() |>
  filter(.metric == "rmse") |>
  slice_min(mean , by = wflow_id) |>
  arrange(mean) |>
  select("Model Type" = wflow_id ,
         "RMSE" = mean ,
         "Standard Error" = std_err ,
         "Num Computations" = n) |>
knitr::kable(digits = c(NA , 2 , 2 , 0))
```
So far both models have an RMSE that is significantly higher than would be useful. Given the possible values for achene count, I would probably want an RMSE in the 15-25 range in order to be confident my model is generating a relatively accurate estimate. However to get that close I would almost certainly need to included data on factors beyond insect presence. For the purposes of this lab I'm aiming to get RMSE under 50.

## Misc
I didn't clarify this in my first progress memo but I'm doing a regression problem.