---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Lena Parnassa"
date: "2/4/2024"

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---



::: {.callout-tip icon=false}

## Github Repo Link

[https://github.com/stat301-2-2024-winter/final-project-2-lena-parnassa](https://github.com/stat301-2-2024-winter/final-project-2-lena-parnassa)

:::
```{r , echo=FALSE}
library(tidyverse)
library(naniar)
library(here)

joined_data <- read_csv(here("data/semi_usable_data.csv"))
```

## Data source

This data is from the lab I'm working in^[https://echinaceaproject.org/]. The specific datasets I'm planning on working with are our [core dataset](https://echinaceaproject.org/datasets/expt1-core/), which includes a few basic metrics recorded annually for each plant in our largest experimental plot, and a datset on [insect damage](https://echinaceaproject.org/datasets/km-aphid-abundance-and-plant-damage-2004-2012/) to the plants in the garden in specific observation years.

## Why this data

I'm an Environmental Science major and the whole reason I wanted to do the Data Science minor was to have a better understanding of the data involved in the field. This project would be a great opportunity for me to apply the understanding I’ve gained so far, and see where the gaps in my knowledge are so I know what I need to improve on.

## Data quality & complexity check
```{r , echo=FALSE}
miss_table <- miss_var_summary(joined_data) |>
  filter(n_miss != 0)
miss_names <- miss_table |>
  pull(variable)

miss_summary <- joined_data |>
  select(all_of(miss_names)) |>
  miss_var_summary() |>
  knitr::kable()

miss_summary
```
## Target Variable Analysis 

## Potential data issues

